### 2024-09-20 cognitive-bias.qmd
20240920 - loss aversion

*Loss aversion bias* is the tendency to avoid losses rather than acquiring equivalent gains.
That is, loss aversion (similar to punishment sensitivity, as opposed to reward sensitivity) is the tendency to weigh losses (or punishment) more heavily than gains (or rewards).
Thus, the target state matters less than the *change* (i.e., gains and/or losses) from one's reference point [e.g., their current roster; @Kahneman2011].
Loss aversion is a key aspect of prospect theory and suggests that individuals tend to be biased in favor of their reference situation and toward making smaller rather than larger changes [@Kahneman2011].
Loss aversion explains why people tend to favor the favoring the status quo; the negative consequences of a change often loom larger than the positive consequences of a change [@Kahneman2011].

The loss aversion ratio indicates the extent to which a person weighs losses greater than gains, and for many people is between 1.5 to 2.5 [@Kahneman2011].
Loss aversion is different from [risk aversion](#sec-cognitiveBiasesRiskAversion).
Loss aversion is exemplified when teams play conservatively so as "not to lose" instead of "to win."
In fantasy football, loss aversion may lead managers to start or hold onto players for too long who were highly drafted yet are underperforming instead of starting a more promising player out of fear of losing potential value from their initial investment.

>>>>

*Loss aversion bias* is the tendency to avoid losses rather than acquiring equivalent gains.
That is, loss aversion (similar to punishment sensitivity, as opposed to reward sensitivity) is the tendency to weigh losses (or punishment) more heavily than gains (or rewards).
Thus, the target state matters less than the *change* (i.e., gains and/or losses) from one's reference point [e.g., their current roster; @Kahneman2011].
Loss aversion is a key aspect of prospect theory and suggests that individuals tend to be biased in favor of their reference situation and toward making smaller rather than larger changes [@Kahneman2011].
Loss aversion explains why people tend to favor the favoring the status quo; the negative consequences of a change often loom larger than the positive consequences of a change [@Kahneman2011].
Aversion to losses and potential threats may be related to humans' survival instinct [@Kahneman2011].
With par as a reference point, this may explain why professional golfers tend to putt more accurately for par than for birdie, even for putts of equivalent distance [@Kahneman2011].

The loss aversion ratio indicates the extent to which a person weighs losses greater than gains, and for many people is around 2 (typically between 1.5 to 2.5; [@Kahneman2011].
That is, people tend to give as much twice the weight to losses as to gains.
Loss aversion is different from [risk aversion](#sec-cognitiveBiasesRiskAversion).
Loss aversion is exemplified when teams play conservatively so as "not to lose" instead of "to win."
In fantasy football, loss aversion may lead managers to start or hold onto players for too long who were highly drafted yet are underperforming instead of starting a more promising player out of fear of losing potential value from their initial investment.
Loss aversion can also influence trade negotiations.
